---
title: "Assignment 3"
author: "Vishal"
date: "2022-10-12"
output: html_document # you can change to other output format if you want
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```

# 1. Exploratory data analysis


## 1.1 (Q1)   

<!-- Include your answers for each question under a second level heading eg ## 1.1 (Q1) -->
**** insert your solutions here *****

```{r}
HawksTail <- Hawks$Tail
HawksTail[1:6]

mean(HawksTail)
median(HawksTail)
```


## 1.2 (Q1)
```{r}
Hawks %>%
  summarise_at(vars(Wing:Weight), c(mean, median), na.rm = T)

Hawks %>%
  summarise(Wing_mean = mean(Wing, na.rm = T), 
            Wing_t_mean = mean(Wing, trim = 0.5, na.rm = T),
            Wing_med = median(Wing, na.rm = T),
            Weight_mean = mean(Weight, na.rm = T),
            Weight_t_mean = mean(Weight, trim = 0.5, na.rm = T),
            Weight_med = median(Weight, na.rm = T))
```
We are calculating mean of 50% trimmed data which essentially means that we are setting aside the top and bottom half and are leaving only a single value in the middle. Which is a median value and that's why mean and median are same when we take a mean with trim value of 0.5.

## 1.2 (Q2)

```{r}
Hawks %>%
  group_by(Species) %>%
  summarise(Wing_mean = mean(Wing, na.rm = T), 
            Wing_t_mean = mean(Wing, trim = 0.5, na.rm = T),
            Wing_med = median(Wing, na.rm = T),
            Weight_mean = mean(Weight, na.rm = T),
            Weight_t_mean = mean(Weight, trim = 0.5, na.rm = T),
            Weight_med = median(Weight, na.rm = T))
```
## 1.3 (Q1)

```{r}
X = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
A = mean(X)

a = 2
b = 3
X_ = (a * X) + b
A_ = mean(X_)

A
A_
```

The mean of `X_` can be defined by `A_ = a * A + b`. That means, we can get the mean of newer vector by multiplying the mean of older vector by value of `a` and adding the value of `b` to the the result.

```{r}
HawksTailMean = mean(HawksTail)
HawksTail_ = (HawksTail * a) + b
HawksTail_Mean = mean(HawksTail_)

HawksTailMean
HawksTail_Mean
```

Hence, it is proved that `A_ = a * A + b`.

## 1.3 (Q2)

```{r}
X = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
varX = var(X)
sdX = sd(X)

a = 4
b = 3

X_ = (a * X) + b
varX_ = var(X_)
sdX_ = sd(X_)

varX
sdX

varX_
sdX_
```
It is seen that variance of a newer vector is square times the value of `a`. 
Whereas the value of standard deviation of the newer vector is `a` times the value of standard deviation of older vector.

```{r}
varHawksTail = var(HawksTail)
sdHawksTail = sd(HawksTail)

a = 2
b = 3
HawksTail_ = (HawksTail * a) + b
varHawksTail_ = var(HawksTail_)
sdHawksTail_ = sd(HawksTail_)

varHawksTail
sdHawksTail

varHawksTail_
sdHawksTail_
```

## 1.4 (Q1)

```{r}
hal <- Hawks$Hallux 
hal <- hal[!is.na(hal)] 
```

```{r}
outlier_val <- 100
num_outliers <- 10
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
```

```{r}
mean(hal)
mean(corrupted_hal)
```
```{r}
num_outliers_vect <- seq(0, 1000)
means_vect <- c()
for(num_outliers in num_outliers_vect){
  corrupted_hal <- c(hal,rep(outlier_val, times = num_outliers))
  means_vect <- c(means_vect, mean(corrupted_hal))
}
```

Sample Median

```{r}
num_outliers_vect <- seq(0, 1000)
medians_vect <- c()
for(num_outliers in num_outliers_vect) {
  corrupted_hal <- c(hal, rep(outlier_val, times = num_outliers))
  medians_vect <- c(medians_vect, median(corrupted_hal))
}

```

## 1.4 (Q2)

Sample trimmed mean
```{r}
num_outliers_vect <- seq(0, 1000)
t_means_vect <- c()
for(num_outliers in num_outliers_vect){
  corrupted_hal <- c(hal,rep(outlier_val, times = num_outliers))
  t_means_vect <- c(t_means_vect, mean(corrupted_hal, trim = 0.1))
}
```

## 1.4 (Q3)

```{r}
df_means_medians <- data.frame(num_outliers = num_outliers_vect, mean = means_vect, 
                               t_mean = t_means_vect, median = medians_vect)

```

```{r}
df_means_medians %>%
  pivot_longer(!num_outliers, names_to = "Estimator", values_to = "Values") %>%
  ggplot(aes(x = num_outliers, color = Estimator, linetype = Estimator, y = Values)) +
  geom_line() +
  xlab("Number of outliers")
```

## 1.5 (Q1)

```{r}
Hawks %>%
  group_by(Species) %>%
  ggplot(aes(x = Species, y = Weight)) +
  geom_boxplot()

```
## 1.5 (Q2)

```{r}
q = c(.25, .50, .75)

Hawks %>%
  group_by(Species) %>%
  summarise(quantile025 = quantile(Weight, probs = q[1], na.rm = T),
            quantile050 = quantile(Weight, probs = q[2], na.rm = T),
            qunatile075 = quantile(Weight, probs = q[3], na.rm = T)) 
```
## 1.5 (Q3)

```{r}
num_outliers <- function(x) {
  
  x<-x[!is.na(x)]
  
  OutliersCount = 0
  q25 = quantile(x, probs = 0.25, na.rm = T)
  q75 = quantile(x, probs = 0.75, na.rm = T)
  
  IQR = q75 - q25
  
  
  for (i in x) {
    if ((i < q25 - (1.5 * IQR)) || (i > q75 + (1.5 * IQR))) {
      OutliersCount = OutliersCount + 1
    }
  }
  return(OutliersCount)
}
num_outliers( c(0, 40,60,185))

```
## 1.5 (Q4)
```{r}
Hawks %>%
  group_by(Species) %>%
  summarise(num_outliers_weight = num_outliers(Weight))
```
## 1.6 (Q1)

```{r}
covWeightWing <- cov(Hawks$Weight, Hawks$Wing, use = "complete.obs")
corrWeightWing <- cor(Hawks$Weight, Hawks$Wing, use = "complete.obs") 

covWeightWing
corrWeightWing
```
## 1.6 (Q2)

```{r}
X <- seq(1, 10, by = 1)
Y <- seq(3, 5.7, by = 0.3)
Y

S <- cov(X, Y)
R <- cor(X, Y)

a = 2
b = 3
c = 5
d = 4
X_ = (a * X) + b
Y_ = (c * Y) + d

S_ = cov(X_, Y_)
R_ = cor(X_, Y_)

S
R
S_
R_
```
The covariance of a newer vector is `a * c` the value of covariance of a older vector. That means `S_ = S * a * c`. Whereas, the correlation stays the same although it becomes negative based on the signs of a and c.

```{r}
a = 2.4
b = 7.1
c = -1
d = 3

X = Hawks$Weight
Y = Hawks$Wing

S = cov(X, Y, use = "complete.obs")
R = cor(X, Y, use = "complete.obs")

X_ = (a * X) + b
Y_ = (c * Y) + d

S_ = cov(X_, Y_, use = "complete.obs")
R_ = cor(X_, Y_, use = "complete.obs")

S
R
S_
R_

```
# 2. Random experiments, events and sample spaces, and the set theory

## 2.1 (Q1)
- **Random Experiment: ** A random experiment is a procedure (real or imagined) which:
1. has a well-defined set of possible outcomes;
2. could (at least in principle) be repeated arbitrarily many times.

- **Event: ** An event is a set (i.e. a collection) of possible outcomes of an experiment.

- **Sample space: ** A sample space is the set of all possible outcomes of interest for a random experiment

## 2.1 (Q2)

**Event: ** One or few of {1, 2, 3, 4, 5, 6}.
**Sample space: ** The whole set of {1, 2, 3, 4, 5, 6}
**Number of different events: ** {1}, {2}, {3}, {4}, {5}, {6} (In case of rolling a single dice! - 6 different events)
Empty set is considered as an event. It is an impossible event.


## 2.2 (Q1)
1. $$A \cup B = {1, 2, 3, 4, 6}$$
   $$A \cup C = {1, 2, 3, 4, 5, 6}$$
   
2. $$A \cap B = {2}$$
   $$A \cap C = {}$$
   
3. $$A \ B = {1, 3}$$
   $$A \ C = {1, 2, 3}$$
   
4. A and B are not disjoint. A and C are disjoint.

5. Yes, $B$ and $A\B$ are disjoint.

6. $${1, 2, 3}, {4, 5, 6$$
   $${1, 2}, {3, 4}, {5,6}$$
   
## 2.2 (Q2)



# 3. Visualisation

## 3 (Q1)

```{r}
Hawks %>%
  group_by(Species) %>%
  ggplot(aes(x = Tail, color = Species)) +
  geom_density() +
  xlab("Tail (mm)") + 
  ylab("Density")
```

## 3 (Q2)
```{r}
Hawks %>%
  group_by(Species) %>%
  ggplot(aes(x = Tail, fill = Species, y = Species)) +
  geom_violin(stat = "ydensity")
```
## 3 (Q3)

```{r}
Hawks %>%
  ggplot(aes(x = Tail, y = Weight, color = Species, shape = Species)) +
  geom_point()
```
## 3 (Q4)

```{r}
Hawks %>%
  ggplot(aes(x = Tail, y = Weight, color = Species)) +
  geom_point() +
  facet_wrap(vars(Species)) +
  geom_smooth() +
  xlab("Tail (mm)") +
  ylab("Weight (gms)")
```

## 3 (Q5)

```{r}
HeaviestHawk <- Hawks %>%
  top_n(1, Weight)
```

```{r}
annotation <- data.frame(
   x = HeaviestHawk$Tail,
   y = HeaviestHawk$Weight,
   label = c("Heaviest Hawk")
)


p <- Hawks %>%
  ggplot(aes(x = Tail, y = Weight, color = Species)) +
  geom_point()

#p + geom_text(data = annotation, aes(x=x, y=y, label=label),                 , 
#           color="red", 
#           size=3 , angle=0, fontface="bold" )

p + annotate("segment", x = annotation$x, xend = annotation$x, y = annotation$y-100, yend = annotation$y, colour = "pink", size=1, alpha=0.6, arrow=arrow())
```
