---
title: "Assignement2"
author: "Vishal"
date: "2022-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data Wrangling
## 1.1 Select and filter
### Q1.

```{r, echo = FALSE}
library(Stat2Data)
library(tidyverse)
library(dplyr)

data("Hawks")

hSF <- Hawks %>%
  filter(Species == "RT" & Weight >= 1000) %>%
  select(Wing, Weight, Tail)

head(hSF)
```
### Q2.
There are 3 variables in the `hSF` called Wing, Weight and Tail. And there are six observations.

## 1.2 The arrange function

```{r}
hSF <- hSF %>%
  arrange(Wing)

head(hSF)
```

## 1.3 Join and rename function
### Q1.
```{r}
hawkSpeciesNameCodes <- data.frame(species_code = c("CH", "RT", "SS"),
                                   species_name_full = c("Cooper's", "Red-tailed", "Sharp-shinned"))

hawkSpeciesNameCodes
```
### Q2.

```{r}
hawksFullName <- Hawks %>%
  rename(species_code = Species) %>%
  left_join(hawkSpeciesNameCodes, on = 'Species') %>%
  select(c(species_name_full, Wing, Weight, Tail)) %>%
  rename(Species = species_name_full)

head(hawksFullName)
```

## 1.4 The mutate function

```{r}
hawksWithBMI <- Hawks %>%
  mutate(bird_BMI = 1000 * Weight / ((Wing) ^ 2)) %>%
  arrange(desc(bird_BMI)) %>%
  select(c(Species, bird_BMI))

head(hawksWithBMI)  

```
## 1.5 Summarise and group-by functions
### Q`.

``` {r}
summary_table <- hawksFullName %>%
  group_by(Species) %>%
  summarise(num_rows = n(),
            mn_wing = mean(Wing),
            md_wing = median(Wing, na.rm = TRUE),
            t_mn_wing = mean(Wing, trim = 0.1),
            b_wt_ratio = max(Wing / Tail, na.rm = T))

summary_table
```

### Q2.
```{r}
hawks_summary <- Hawks %>%
  select(c(Species, Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Crop)) %>%
  group_by(Species) %>%
  summarise(across(everything(), ~sum(is.na(.))))

hawks_summary
  
```

# 2. Tidy data and iteration
## 2.1 Missing data and iteration

### Q1. 
We are using a function called `map_dbl` inside our custom function which is a part of `tidyverse` library. So, we need `tidyverse` library to run our custom function.

### Q2.
```{r}
impute_by_median <- function(x) {
  md <- median(x, na.rm = T)
  impute_f <- function(z) {
    if(is.na(z)) {
      return(md)
    } else {
      return(z)
    }
  }
  return(map_dbl(x, impute_f))
}

V <- c(1, 2, NA, 4)
impute_by_median(V)
```
### Q3.

```{r}
x <- seq(0, 10, by = 0.1)
y <- seq(1, 51, by = 0.5)

df_xy <- data.frame(x, y)
head(df_xy)

df_xy %>% mutate(z=map2_dbl(x,y,~.x+.y)) %>% head(5)
```
### Q4.
```{r}
sometimes_missing <- function(index, value) {
  if (index %% 5 == 0) {
    return(NA)
  } else {
    return(value)
  }
}

sometimes_missing(14, 25)
sometimes_missing(15, 25)


df_xy_missing <- df_xy %>%
  mutate(y = map2_dbl(row_number(), y, sometimes_missing))
df_xy_missing <- df_xy %>%
  mutate(y = map2_dbl(.x = row_number(), .y = y, sometimes_missing))

head(df_xy_missing, 10)
```

### Q5.
```{r}
df_xy_imputed <- df_xy_missing %>%
  mutate(y = impute_by_median(y))
head(df_xy_imputed)

```
## 2.2 Tidying data with pivot functions
```{r}
library(readxl)

library(ggplot2)
folder_path <- "D://Vishal//Bristol//Stats//"
file_name <- "HockeyLeague.xlsx"
file_path <- paste(folder_path, file_name, sep = "")
wins_data_frame <- read_excel(file_path, sheet = "Wins")
head(wins_data_frame)
```

### Q1. 
```{r}
wins_tidy <- wins_data_frame %>%
  pivot_longer(!...1) %>%
  rename(Year = name, Team = ...1) %>%
  separate(value, c("Wins", "Total"), sep = " of ")

wins_tidy[, 2:4] <- sapply(wins_tidy[, 2:4], as.integer)
head(wins_tidy)
```

### Q2. 
```{r}
losses_data_frame <- read_excel(file_path, sheet = "Losses")
head(as.data.frame(losses_data_frame))
```

```{r}
losses_tidy <- losses_data_frame %>%
  pivot_longer(!...1) %>%
  rename(Year = name, Team = ...1) %>%
  separate(value, c("Losses", "Total"), sep = " of ")
head(losses_tidy)
```

### Q3.
```{r}
hockey_df <- merge(wins_tidy, losses_tidy, by = c('Team', 'Year')) %>%
  select(-c(Total.y)) %>%
  rename(Total = Total.x) 

hockey_df[, 2:5] <- sapply(hockey_df[, 2:5], as.integer)

hockey_df <- hockey_df %>%
  mutate(Draws = Total - Wins - Losses,
         Wins_rt = Wins / Total,
         Losses_rt = Losses / Total,
         Draws_rt = Draws / Total
         )

head(hockey_df)
```

### Q4.
```{r}
summary_table <- hockey_df %>%
  select(c(Team, Wins_rt, Losses_rt, Draws_rt)) %>%
  group_by(Team) %>%
  summarise(across(everything(), list(md = median, mn = mean))) %>%
  arrange(desc(Wins_rt_md))

summary_table 
```

# Visulisation

### Q1.
```{r}
wins_tidy %>%
  filter(Team == "Ducks") %>%
  ggplot(aes(x = as.numeric(Wins))) +
  geom_histogram(binwidth = 3) +
  xlab("Wins") + 
  ylab("Count")
```
### Q2.
```{r}
wins_tidy %>%
  filter(Team == "Ducks") %>%
  ggplot(aes(x = as.numeric(Wins))) +
  geom_density(adjust = 0.5) +
  xlab("Wins") + 
  ylab("Count")
```

```{r}
wins_tidy %>%
  filter(Team == "Ducks") %>%
  ggplot(aes(x = as.numeric(Wins))) +
  geom_density(adjust = 2) +
  xlab("Wins") + 
  ylab("Count")
```

Explanation: adjust is used the control the bandwidth while still using the bandwidth estimator.

### Q3.
```{r}
wins_teams <- wins_tidy %>%
  select(-Total) %>%
  pivot_wider(names_from = Team, values_from = Wins)

wins_teams[, 1:9] <- sapply(wins_teams[, 1:9], as.integer)

wins_teams %>%
  ggplot(aes(x = Ducks, y = Eagles)) +
  geom_point()
```