---
title: "Assignment 10"
author: "Vishal"
date: "2022-12-07"
output: pdf_document
---

```{r setup, include=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Linear discriminant analysis

## Q1.

Probabilistic LDA is a generative model which assumes that given data samples are generated from a distribution (e.g. Gaussian distribution). We need to find the parameters of model which best describe the training data. 
```{r}
library(Stat2Data) 
library(tidyverse)
library(MASS)
library(ggplot2)
library(glmnet)
library(QSARdata)
data(Hawks)
```

```{r}
hawks_total <- Hawks %>% 
  dplyr::select(Weight, Wing, Hallux, Tail, Species) %>% 
  filter(Species =='SS' | Species =='CH') %>% drop_na() %>% 
  mutate(Species = as.numeric(Species =='SS'))

head(hawks_total)
```

```{r}
num_total <- hawks_total %>% nrow() 
num_train <- floor(num_total * 0.6) 
num_test <- num_total-num_train 
set.seed(0) 
test_inds <- sample(seq(num_total), num_test) 
train_inds <- setdiff(seq(num_total), test_inds) 
hawks_train <- hawks_total %>% 
  filter(row_number() %in% train_inds) 

hawks_test <- hawks_total %>%  
  filter(row_number() %in% test_inds) 
```


```{r}
hawks_train_x <- hawks_train %>% 
  dplyr::select(-Species)
hawks_train_y <- hawks_train %>% 
  pull(Species)

hawks_test_x <- hawks_test %>% 
  dplyr::select(-Species)
hawks_test_y <- hawks_test %>% 
  pull(Species)
```

```{r}
lda_model <- MASS::lda(Species ~ ., hawks_train)
```

## Q2.

```{r}
lda_train_predicted_y <- predict(lda_model, hawks_train_x)$class %>%
  as.character() %>% as.numeric()

lda_train_error <- mean(abs(lda_train_predicted_y - hawks_train_y))
lda_train_error
```
```{r}
lda_test_predicted_y <- predict(lda_model, hawks_test_x)$class %>%
  as.character() %>% as.numeric()

lda_test_error <- mean(abs(lda_test_predicted_y - hawks_test_y))
lda_test_error
```

## Q3. 

```{r}
data("iris")

iris_total <- iris %>%  drop_na() %>% 
  mutate(Species = as.numeric(Species))

head(iris_total)
```

```{r}
num_total <- iris_total %>% nrow() 
num_train <- floor(num_total * 0.6) 
num_test <- num_total - num_train 
set.seed(0) 
test_inds <- sample(seq(num_total), num_test) 
train_inds <- setdiff(seq(num_total), test_inds) 
iris_train <- iris_total %>% 
  filter(row_number() %in% train_inds) 

iris_test <- iris_total %>%  filter(row_number() %in% test_inds) 
```

```{r}
iris_train_x <- iris_train %>% dplyr::select(-Species)
iris_train_y <- iris_train %>% pull(Species)

iris_test_x <- iris_test %>% dplyr::select(-Species)
iris_test_y <- iris_test %>% pull(Species)
```

```{r}
lda_model <- MASS::lda(Species ~ ., iris_train)
```

```{r}
lda_train_predicted_y <- predict(lda_model, iris_train_x)$class %>%
  as.character() %>% as.numeric()

lda_train_error <- mean(abs(lda_train_predicted_y - iris_train_y))
lda_train_error
```
```{r}
lda_test_predicted_y <- predict(lda_model, iris_test_x)$class %>%
  as.character() %>% as.numeric()

lda_test_error <- mean(abs(lda_test_predicted_y - iris_test_y))
lda_test_error
```
# 2. Logistic regression

## Q1. \
We know that, in Logistic regression target variable follows Bernoulli distribution. Probability mass function of Bernoulli distribution is \
$$\mathbb{P}(y) = p^y * (1 - p)^{(1 - y)}$$\
We also know that, Generalized Linear Model is of the form:\
$$y = w^Tx + \epsilon \\
i.e. f(E(y)) = w^Tx$$ \

Logit function is a link function in this kind of Generalized Linear Models. Logit function is defined as : log(odds) i.e. \
**odds = (success/failure)**. \

$$odds = \frac{p}{(1 - p)} \\
log(odds) = log (\frac{p}{1 - p})\\
\Rightarrow f(p) = log(\frac{p}{1 - p})$$

For Bernoulli distribution $E(y) = p$. Hence equation becomes, $f(E(y)) = f(p)$ \

$$f(p) = f(E(y)) \\
log(frac{p}{1 - p}) = w ^ T x \\
\text {Taking exponential}, \\
\frac{p}{1 - p} = exp(w^Tx) \\
p = (1 - p)exp(w^Tx) \\
p = exp(w^Tx) - p.exp(w^Tx) \\
p + p.exp(w^Tx) = exp(w^Tx) \\
p(1 + exp(w^Tx)) = exp(w^Tx) \\
p = \frac{exp(w^Tx)}{1 + exp(w^Tx)} \\
\text{OR} \\
p = \frac{1}{1 + exp(-w^Tx)}$$

## Q2.
```{r, results='hide', fig.keep='all'}
sigmoid <- function(z) {
  S_z <- 1 / (1 + exp(-z))
  return(S_z)
}

z <- seq(-10, 10, 0.1)

temp_df <- data.frame(z = seq(-10, 10, 0.1)) %>%
  mutate(S_z = map_dbl(z, sigmoid))

ggplot(temp_df, aes(x = z, y = S_z)) +
  geom_line(color = "blue") +
  labs(y = 'S(z)')
  theme_bw()
```

## Q3.
```{r}
logistic_model <- glmnet(x = hawks_train_x %>% as.matrix(), 
                         y = hawks_train_y, 
                         family = 'gaussian',
                         alpha = 0,
                         lambda = 0)
```

```{r}
logistic_train_predicted_y <- predict(logistic_model, hawks_train_x %>%
                                        as.matrix(),
                                      type = 'class') %>%
  as.integer()
logistic_train_error <- mean(abs(logistic_train_predicted_y - hawks_train_y))

logistic_train_error
```

```{r}
logistic_test_predicted_y <- predict(logistic_model, hawks_test_x %>%
                                       as.matrix(),
                                     type = 'class') %>%
  as.integer()

logistic_test_error <- mean(abs(logistic_test_predicted_y - hawks_test_y))

logistic_test_error
```

# 3. Basic concepts in regularisation
## Q1.
### **1. Hyper-parameter:**\
Hyper-parameters are parameters whose values control the learning process and determine the values of model parameters that a learning algorithm ends up learning. They are used to to control the learning process and the model parameters that result from it.\
Learning rate & number of epochs are examples of hyper-parameters.\

### **2. Validation data:**\
A validation dataset is a sample of data held back from training your model that is used to give an estimate of model skill while tuning model’s hyper-parameters. \
It is used to give an unbiased estimate of the skill of the final tuned model when comparing or selecting between final models. \

### **3. The train-validation-test split:**\
The train-test-validation-test split is a technique to evaluate the performance of a machine learning model. A given dataset is divided into three subsets for different purposes.\
Train datset is used to train the model initially. \
Validation dataset is used to provide an unbiased evaluation of a model fitted on the training dataset while tuning hyper-parameters.\
Whereas, test dataset is used to provide an unbiased evaluation of a final model fitted on the training dataset.\
  

## Q2.
**$l_2$ norm:**\
The Euclidean norm of a vector which is a point on a line, surface, or hyper-surface may be interpreted geometrically as the distance between this point and the origin.\
$$||x||_2 = \sqrt{x_1^2 + x_2^2 + x_3^2 + ... +x_n^2}$$
**$l_1 norm:$**\
$l_1$ norm is also known as "Manhattan Distance or Taxicab norm." It is the sum of the magnitudes of the vectors in a space i.e. sum of absolute difference of the components of vectors. \
$$||x||_1 = |x_1| + |x_2| + |x_3| + ... + |x_n|$$

## Q3.
**Lasso Regression Regularization:**\
It is a shrinkage technique. It stands for Least Absolute Shrinkage and Selection Operator. It is used over regression methods for a more accurate prediction. This model uses shrinkage.  Shrinkage is where data values are shrunk towards a central point as the mean. The lasso procedure encourages simple, sparse models (i.e. models with fewer parameters).\

$$L_{lasso}(\hat{\beta}) = \sum_{i=1}^{n}(y_i - x_i\hat{\beta})^2 + \lambda \sum_{j = 1}^{m}|\hat{\beta_j}|$$
**Ridge Regression:**\
Similar to the lasso regression, ridge regression puts a similar constraint on the coefficients by introducing a penalty factor. However, while lasso regression takes the magnitude of the coefficients, ridge regression takes the square.\
$$L_{hridge}(\hat{\beta}) = \sum_{i = 1}^{n}(y_i - x_i \hat{\beta})^2 + \lambda \sum_{j = 1}^{m} w_j \hat{\beta_j}^2$$
Ridge regression is also referred to as L2 Regularization.

# 4. An investigation into ridge regression for high-dimensional regression

```{r}
data(MeltingPoint)
```

```{r}
mp_data_total <- MP_Descriptors %>% 
  mutate(melting_pt=MP_Outcome)
```

There are **203** variables in the **mp_data_total** data frame. And there are **4401** examples.\

## Q2.
```{r}
spec = c(train = .5, test = .25, val = .25)

g = sample(cut(
  seq(nrow(mp_data_total)), 
  nrow(mp_data_total) * cumsum(c(0, spec)),
  labels = names(spec)
))

res <- split(mp_data_total, g)

mp_data_train <- res$train
mp_data_test <- res$test
mp_data_val <- res$val
```

## Q3.
```{r}
val_error_calculator <- function(train_data, val_data, ld) {
  
  train_x <- train_data %>% 
    dplyr::select(-melting_pt)
  
  train_y <- train_data %>% 
    pull(melting_pt)
  
  val_x <- val_data %>% 
    dplyr::select(-melting_pt)
  
  val_y <- val_data %>% 
    pull(melting_pt)
  
  logistic_model <- glmnet(x = train_x %>% as.matrix(), 
                         y = train_y, 
                         family = 'gaussian',
                         alpha = 1,
                         lambda = ld)
  
  logistic_val_predicted_y <- predict(logistic_model, val_x %>%
                                         as.matrix(),
                                       type = 'class') %>%
    as.integer()

  logistic_val_error <- mean(abs(logistic_val_predicted_y - val_y))
  
  return(logistic_val_error)
}

val_error_calculator(mp_data_train, mp_data_val, 0.1)

```
## Q4.
```{r}
mySeq <- seq(0, 70)
lambdas <- c()
for (i in mySeq) {
  temp <- (1.25 ^ i) * (10 ^ -5)
  lambdas <- c(lambdas, temp)
}

lambdas
```

## Q5.
```{r}
df <- data.frame(lambda = lambdas) %>%
  mutate(mse = map_dbl(.x = lambda, .f = ~val_error_calculator(mp_data_train, mp_data_test, .x)))
```

## Q6.
```{r}
ggplot(df, aes(x = lambda, y = mse)) +
  geom_point() +
  scale_x_log10() +
  theme_bw()
```
## Q7.
```{r}
df[which.min(df$mse), ]

val_error_calculator(mp_data_train, mp_data_test, 0.0385186)
```













