---
title: "Assignment07"
author: "Vishal"
date: "2022-11-16"
output: pdf_document
---

<<<<<<< HEAD
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# 1. Maximum likelihood estimates

```{r}
library(Stat2Data) 
library(tidyverse)
data("Hawks")
```

# 1. Maximum likelihood estimates

## 1.1 (Q1)
```{r}
RedTailedDf <- Hawks %>%
  filter(Species == "RT") %>%
  select(Weight, Tail, Wing)

head(RedTailedDf)
```

## 1.1 (Q2)
```{r}
RedTailedVector <- RedTailedDf$Tail
mu_RedTailedVector <- mean(RedTailedVector)
sigma_RedTailedVector <- sd(RedTailedVector)

mle <- sum((RedTailedVector - mu_RedTailedVector) ^ 2) / length(RedTailedVector)

guassian_df <- data.frame(Length = RedTailedVector) %>%
  mutate(pdf = map_dbl(Length, ~dnorm(.x, mu_RedTailedVector, sigma_RedTailedVector)))

colors <- c("MLE density" = "red", "Kernel density" = "blue")

p <- ggplot() + labs(y = "Tail Length (mm)") + theme_bw() +
  geom_density(data = guassian_df, aes(Length, color = "Kernel density"))

p +
  geom_line(data = guassian_df, aes(Length, pdf, color = "MLE density")) +
  scale_color_manual(values = colors)
```

## 1.1 (Q3)

```{r}
mle_calc <- function(X, X_bar, len) {
  mleVal = sum((X - X_bar) ^ 2) / len
  return(mleVal)
}
RedTailedDfVector <- RedTailedDf$Tail
n <- length(RedTailedDfVector)
mu_mle <- mean(RedTailedDfVector)
sigma_mle <- sd(RedTailedDfVector) * sqrt((n - 1) / n)

mu_mle
sigma_mle
```

## 1.1 (Q3)
```{r}
density_df <- data.frame(Length = RedTailedDfVector) %>%
  mutate(pdf = map_dbl(.x = Length, .f = ~dnorm(.x, mean = mu_mle, sd = sigma_mle)))


colors <- c("MLE Density" = "red", "Kernel Density" = "blue")

density_df %>%
  ggplot() +
  geom_line(aes(x = Length, y = pdf, color = "MLE Density")) +
  geom_density(aes(x = Length, color = "Kernel Density")) +
  scale_color_manual(values = colors) +
  labs(y = "Tail length (mm)") +
  theme_bw()
```

## 1.2 (Q1)
```{r}
df <- data.frame(sampleSize = rep(seq(5, 100, 5), each = 1000)) %>%
  mutate(sample = map(.x = sampleSize, ~(rnorm(.x, 1, 3))),
         mu = map_dbl(.x = sample, .f = mean),
         v_mle = map_dbl(.x = sample, .f = sd),
         v_u = map_dbl(.x = sample, .f = var))

d <- df %>%
  group_by(sampleSize) %>%
  summarise(across(c(v_mle, v_u), mean)) %>%
  pivot_longer(!sampleSize, names_to = "attr", values_to = "Values")

head(d)
ggplot(d, aes(x = sampleSize, y = Values, color = attr)) +
  geom_line()
```
## 1.2 (Q2)
```{r}
df <- data.frame(sampleSize = rep(seq(5, 100, 5), each = 1000)) %>%
  mutate(sample = map(.x = sampleSize, ~(rnorm(.x, 1, 3))),
         mu = map_dbl(.x = sample, .f = mean),
         v_mle = map_dbl(.x = sample, .f = sd),
         v_u = map_dbl(.x = sample, .f = var),
         v_u_sqrt = map_dbl(.x = v_u, .f = sqrt))

d <- df %>%
  group_by(sampleSize) %>%
  summarise(across(c(v_mle, v_u_sqrt), mean)) %>%
  pivot_longer(!sampleSize, names_to = "attr", values_to = "Values")

head(d)
ggplot(d, aes(x = sampleSize, y = Values, color = attr)) +
  geom_line()
```
\
As it can be seen in the above graph, $\sqrt{Vu}$ is not an unbiased estimator of for $\sigma_0$. Calculating the sample variance treats the sample mean as the true mean, where there's an uncertainty about the true mean. This is why square root of variance is not a good estimator.

## 1.3 Maximum likelihood estimation with the Poisson distribution

## 1.3 (Q1)
$$f(x)=\frac{{\lambda}^{x}{e}^{-\lambda}}{x!} x=0, 1, 2,$$
$$L(\lambda)=\prod_{i=1}^{n}\frac{{\lambda}^{{x}_{i}}{e}^{-\lambda}}{{x}_{i}!} = {e}^{-n\lambda} \frac{{\lambda}^{\sum_{1}^{n}{x}_{i}}}{\prod_{i=1}^{n}{x}_{i}}$$
$$lnL(\lambda)=-n\lambda+\sum_{1}^{n}{x}_{i}ln(\lambda)-ln\left(\prod_{i=1}^{n}{x}_{i}\right)$$
$$\frac{dlnL(\lambda)}{dp}=-n+\sum_{1}^{n}{x}_{i}\frac{1}{\lambda}$$

$$\hat{\lambda}=\frac{\sum_{i=1}^{n}{x}_{i}}{n}$$

## 1.3 (Q2)
Suppose that $X = (X_1, X_2, ..., X_n)$ are iid observations from a Poisson distribution with unknown parameter $\lambda$. The likelihood function is\
$$\begin{aligned} L(\lambda) =\prod\limits_{i=1}^{n} f\left(x_{i} ; \lambda\right) =\prod\limits_{i=1}^{n} \dfrac{\lambda^{x_{i}} e^{-\lambda}}{x_{i} !}  =\dfrac{\lambda^{\sum_i x_{i}} e^{-n \lambda}}{x_{1} ! x_{2} ! \cdots x_{n} !} \end{aligned}$$
The corresponding loglikelihood function is\
$$\sum\limits_{i=1}^{n} x_i\log\lambda-n\lambda-\sum\limits_{i=1}^{n} x_i!$$
And the MLE for $\lambda$ can then be found by maximizing either of these with respect to $\lambda$. Setting the first derivative equal to 0 gives the solution:
$$\hat{\lambda}=\sum\limits_{i=1}^{n} \dfrac{x_i}{n}$$

Thus, for a Poisson sample, the MLE for $\lambda$ is just the sample mean.

## 1.3 (Q3)
```{r}
dfPoisson <- data.frame(data = rpois(1000, 0.5))

dfPoisson %>%
  ggplot(aes(x = data)) +
  geom_histogram(bins = 30) +
  labs(y = "Count", x = "data")

```
# 1.3 (Q4)
```{r}
df_VonBortkiewicz <- read.csv("VonBortkiewicz.csv")
head(df_VonBortkiewicz)
```

```{r}
# likelihood of single data point
#likelihood <- data.frame(data = dpois(df_VonBortkiewicz$fatalities[1], seq(20)))

likelihood <- dpois(df_VonBortkiewicz$fatalities[1], seq(20))

lh_single <- data.frame(x = seq(0, 9.5, 0.5), likelihood = likelihood)

head(lh_single)

lh_single %>%
  ggplot(aes(x = x, y = likelihood)) +
  geom_point(size = 3, color = "blue") +
  labs(x = "Lambda", y = "Likelihood") +
  theme_bw()
#ggplot(horseFatilities, aes(x = data)) +
#  geom_histogram(bins = 100)

## Log likelihood of single data point
log_likelihood <- dpois(df_VonBortkiewicz$fatalities[1], seq(20), log = T)
llh_single <- data.frame(x = seq(0, 9.5, 0.5), log_like = log_likelihood)

llh_single %>%
  ggplot(aes(x = x, y = log_like)) +
  geom_point(size = 3, color = "red") +
  labs(x = "Lambda", y = "Log likelihood") +
  theme_bw()

```
```{r}
llh_poisson <- function(lambda, y) {
  llh <- sum(dpois(y, lambda, log = T))
  return(llh)
}
```

```{r}
samples <- data.frame(sample_size = rep(seq(5, 100, 5), each = 1000)) %>%#(temp = seq(5, 100, 5)) %>%
  mutate(sample_X = map(sample_size, ~rnorm(.x, 1, 3)),
         X_bar = map_dbl(sample_X, mean),
         V_mle = map_dbl(sample_X, ~mle_calc(.x, mean(.x), length(.x))),
         V_u = map_dbl(sample_X, ~mle_calc(.x, mean(.x), (length(.x) - 1))))

head(samples)
```

```{r}
lambdas <- seq(0, 9.5, 0.5)
```

```{r}
ll <- sapply(lambdas, function(x) {llh_poisson(x, df_VonBortkiewicz$fatalities)})

df <- data.frame(ll = ll, lambda = lambdas)
```

```{r}
df %>%
  ggplot(aes(x = lambda, y = ll)) +
  geom_point(size = 4, color = "blue") +
  labs(x = "Lambda", y = "Log likelihood") +
  geom_vline(xintercept = lambdas[which.max(ll)], color = "red", size = 2) +
  theme_bw()
```
From above graph, we can see that $\lambda = 0.5$.

```{r}
df_VonBortkiewicz_I <- df_VonBortkiewicz %>%
  filter(corps == 'I')

noFatilities <- data.frame(data = dpois(df_VonBortkiewicz_I$fatalities, 0.5))
noFatilities %>%
  ggplot(aes(x = data)) +
  geom_histogram(bins = 10)
```

## 1.4 Maximum likelihood estimation for the exponential distribution

## 1.4 (Q1)
$$\mathcal{L}(\lambda,x_1,\dots,x_n)=\prod_{i=1}^n f(x_i,\lambda)=\prod_{i=1}^n \lambda e^{-\lambda x}=\lambda^ne^{-\lambda\sum_{i=1}^nx_i}$$
Maximum likelihood estimator:\
$$\frac{d\ln\left(\mathcal{L}(\lambda,x_1,\dots,x_n)\right)}{d\lambda}\overset{!}{=}0$$
$$\frac{d\ln\left(\mathcal{L}(\lambda,x_1,\dots,x_n)\right)}{d\lambda} =\frac{d\ln\left(\lambda^ne^{-\lambda\sum_{i=1}^nx_i}\right)}{d\lambda} \\
= \frac{d\ln\left(n\ln(\lambda)-\lambda\sum_{i=1}^n x_i\right)}{d\lambda} \\
= \frac{n}{\lambda}-\sum_{i=1}^n x_i$$

Which equals to,\
$$\lambda = \frac{n}{\sum\limits_{i=1}^n x_i}$$
## 1.4 (Q2)
```{r}
df_CustomerPurchases <- read.csv("CustomerPurchase.csv")

df_CustomerPurchases$lead <- lead(df_CustomerPurchases$Time)
df_CustomerPurchases$time_diffs <- df_CustomerPurchases$lead - df_CustomerPurchases$Time
df_CustomerPurchases <- select(df_CustomerPurchases, -lead)

head(df_CustomerPurchases)
```
## 1.4 (Q3)
```{r}
exp_mle <- 1 / mean(df_CustomerPurchases$time_diffs, na.rm = T)
exp_mle
```

## 1.4 (Q4)
```{r}
moreThanMin <- filter(df_CustomerPurchases, time_diffs > 60)
head(moreThanMin)

dist <- data.frame(data = pexp(df_CustomerPurchases$time_diffs, rate = 0.02)) %>%
  mutate(time_diff = df_CustomerPurchases$time_diffs)

ggplot(dist, aes(x = time_diff, y = data)) +
  geom_point()
```

# 2. Confidence intervals
## 2.1 Student's t-confidence intervals#

## 2.1 (Q1)

1. If the sample mean increases, width of confidence interval will not change.\
2. If the sample standard deviation was higher, the width of confidence interval would have increased.\
3. The width of confidence interval decreases when the sample size increases.\

## 2.1 (Q2)

```{r}
HawksVector <- Hawks %>%
  filter(Species == "RT") %>%
  pull(Weight)

HawksVector <- HawksVector[!is.na(HawksVector)]
head(HawksVector)
```

```{r}
alpha <- 0.01
sample_size <- length(HawksVector)
sample_mean <- mean(HawksVector)
sample_sd <- sd(HawksVector)
t <- qt(1 - alpha / 2, df = sample_size - 1)
# confidence interval
confidence_interval_l <- sample_mean - t * sample_sd / sqrt(sample_size)
confidence_interval_u <- sample_mean + t * sample_sd / sqrt(sample_size)
confidence_interval <- c(confidence_interval_l, confidence_interval_u)
confidence_interval
```


```{r}
DF <- length(HawksVector) - 1
alpha <- 0.01
t = qt(1 - alpha / 2, DF)
t

t.test(HawksVector, conf.level = 0.99)
```
```{r}

tempDf <- data.frame(x = HawksVector) %>%
  mutate(pdf = map_dbl(.x = x, ~dnorm(.x, mean(HawksVector), sd(HawksVector))))

ggplot(data = tempDf, aes(x = x)) +
  geom_density() 

ggplot(tempDf, aes(sample = x)) + 
  stat_qq() +
  stat_qq_line()
```

## 2.2 Investigating coverage for Student’s t intervals
## 2.2 (Q1)

```{r}
student_t_confidence_interval <- function(sample, confidence_level) {
  sample <- sample[!is.na(sample)] # remove any missing values
  n <- length(sample) # compute sample size
  mu_est <- mean(sample) # compute sample mean
  sig_est <- sd(sample) # compute sample sd
  alpha = 1 - confidence_level # alpha from gamma
  t <- qt(1 - alpha / 2, df = n - 1) # get student t quantile
  l = mu_est - (t / sqrt(n)) * sig_est # lower
  u = mu_est + (t / sqrt(n)) * sig_est # upper
  return(c(l, u))
}
```

```{r}
num_trials <- 100000
sample_size <- 30
mu_0 <- 1
sigma_0 <- 3
alpha <- seq(0.01, 0.09, 0.01)
set.seed(0) # set random seed for reproducibility


probEstimate <- function(num_trials, sample_size, mu_0, alpha) {
  single_alpha_coverage_simulation_df <- data.frame(trial = seq(num_trials)) %>%
    mutate(sample = map(.x =  trial,.f = ~rnorm(n = sample_size, mean = mu_0, sd = sigma_0))) %>%
    mutate(ci_interval = map(.x = sample, .f = ~student_t_confidence_interval(.x, 1 - alpha)))%>%
    mutate(cover = map_lgl(.x = ci_interval, .f = ~((min(.x) <= mu_0) & (max(.x) >= mu_0))))%>%
    mutate(ci_length = map_dbl(.x = ci_interval, .f = ~(max(.x) - min(.x))))
  
  
  val <- single_alpha_coverage_simulation_df %>%
    pull(cover) %>%
    mean()
  return(val)
}

lu_vec <- c()
gamma_vec <- c()

for(i in alpha) {
  g <- probEstimate(num_trials, sample_size, mu_0, i)
  g
}
```
# 3. One sample hypothesis testing
## 3.1 One sample t-test on penguins data

## 3.1 (Q1)
```{r}
library(palmerpenguins)
data(package = 'palmerpenguins')
head(penguins)
```
```{r}
bill_adelie <- penguins %>%
  filter(species == "Adelie") %>%
  pull(bill_length_mm)

t.test(bill_adelie, mu = 40, conf.level = 0.01)
```

## 3.2 Implementing a one-sample t-test
## 3.2 (Q1)
```{r}
p_val_calc <- function(x, mu) {
  mn <- mean(x, na.rm = T)
  t = (mn - mu) / (sd(x, na.rm = T) / sqrt(length(x)))
  p = 2 * (1 - pt(abs(t), df = length(x) - 1))
  return(p)
}

p_val_calc(bill_adelie, 40)
```
